---
title: "Exercise 3"
output: html_document
---

#Exercise 3.

```{r include=FALSE}
# read the data into memory and access dplyr
library(dplyr)
alc <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt", sep=",", header=TRUE)


```



##Task 2.

Names and description of the variables.

```{r}
glimpse(alc)
summary(alc)


```


##Task 3.    

The chosen variables are:

sex - student's sex (binary: 'F' - female or 'M' - male)     
higher - wants to take higher education (binary: yes or no)      
failures - number of past class failures (numeric: n if 1<=n<3, else 4)      
G3 - final grade (numeric: from 0 to 20, output target)     

Hypothesises:
There are more male high users than females.  
Those who want to get a higher education drink less.
Alcohol abuse affects the likelihood of passing the class.
G3 is lower among high users.

##Task 4.
Numerically and graphically explore the distributions of your chosen variables and their relationships with alcohol consumption (use for example cross-tabulations, bar plots and box plots). Comment on your findings and compare the results of your exploration to your previously stated hypotheses. 
The factors sex, goals,failures and will to get a high education has been individually compared to alcohol consumption with cross-tabulations.
The relation between alcohol consumption and final grade number has been analyzed with the help of box plots.

```{r}

table(high_use=alc$high_use,sex=alc$sex) %>% prop.table %>% addmargins

table(high_use=alc$high_use,higher_education=alc$higher) %>% prop.table %>% addmargins

table(failures=alc$failures,high_use=alc$high_use) %>% prop.table %>% addmargins

# library
library(ggplot2)
# A really basic boxplot.
ggplot(alc, aes(x=as.factor(high_use), y=G3)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + 
    xlab("High user")
 

```

The results show that there are more male heavy drinkers than female and that those who want a high education drinks less than those with not such high ambitions. There is not a clear pattern to be found between the drinking habits and the failures in class and the final grade.

##Task 5.

The relationships between the previous variables explained with logistical regression.

```{r}
m <- glm(high_use ~ sex+failures + higher+G3, data = alc, family = "binomial")

summary(m)
#OR:t
cbind(exp(coef(m)),exp(confint(m)))

```

Males posess a ~2.3 times higher risk of being high users compared to the females. The sex and the number of failures in class are related to high alcohol consumptio, the others are not.

##Task 6.

Provide a 2x2 cross tabulation between the variables with a statistical relationship with high/low alcohol consumption.
```{r}
# predict() the probability of high_use
probabilities <- predict(m, type = "response")

# add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)
alc <- mutate(alc, prediction = probability > 0.5)

g <- ggplot(alc, aes(x = probability, y = high_use,col = prediction))
g+geom_point()

# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table %>% addmargins

```
This model predicts more of not high users than high users which means it's not reliably accurate.
